---
title: "An introduction of data.vcf"
author: "liqg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The data.vcf is a fast and simple tool to manipulate VCF in R. 

- Fast: rewriting `read_lines` as an alternative of the R's `readLines` for fast plain text or gz compressed file reading.
- Simple: all items in INFO and FORMAT are represented as an uniform format, a data.table including `varid-key-value`. For example, the INFO is converted as a data table of 3 columns: `varid`(internal IDs for variants), `k` (keys of an INFO) and `v` (values of an INFO); FORMAT and individual columns are also converted to the similar format, a data.table including `varid`(indexs of variants), `k` (keys of FORMAT), sample1 and so on. This table-like `varid-key-value` structure can be easily manipulated in R. All the time-consuming string manipulations in this reformating process are also writen in C++ using Rcpp for fast speed.

### Open a VCF file
```{r}
library(data.vcf)
vcf <- open_vcf(system.file("test.vcf.gz", package="data.vcf"))
names(vcf)
```
`open_vcf` will return an environment object including the meta lines `meta`, header colnames `header`, and a file connection `con` which is used for lines reading.

## Read and convert variants
```{r}
lines <- read_body(vcf, n=100) # read 100 lines
variants <- reformat_body(lines)
names(variants)
```

`C7` is the columns 1-7 of the vcf file
```{r}
knitr::kable(head(variants$C7))
```

`INFO` is a data.table for INFO
```{r}
knitr::kable(head(variants$INFO))
```

`FORMAT` is a data.table for FORMAT and samples
```{r}
knitr::kable(head(variants$FORMAT))
```

### Revert formated variants for outputing new vcf file
```{r}
vr <- tobody(variants)
```

### output a new vcf file
```{r, eval=FALSE}
newvcf <- "new.vcf"
write(vcf$meta, file=newvcf)
write(paste0("#", paste(vcf$header, collapse="\t")), file=newvcf, append=TRUE)
fwrite(vr, file=newvcf, append=T)
```

